Manuscript Preparation

### 2018-06-18 Houston/src/manuscript_preparation.Rmd
Random forest and tuning

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
options(scipen = 1, digits = 2)
require(dplyr)
require(ggplot2)
require(RcppRoll)
require(markovchain) # FIX: markovchain seems not compatible with R3.5.0
require(reshape)
require(survival)
require(caret)
require(e1071)
# knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/prepare_violation.R")
# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/roll_over.R")
#knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN

@@

num.days.lm <- lm(0*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)
```
Linear regression on num of deaths:

@@

@@

@@

# save the computationally costly svm model
#save(svm.model, file="./Houston/data/svm_radial.RData")

plot(svm.model, labeled.data.lost.train)

compareTable <- table (labeled.data.lost.train$severe, predict(svm.model))  # tabulate
mean(labeled.data.lost.train$severe != predict(svm)) # false rate

pred.train <- predict(svm.model, labeled.data.lost.train)
mean(pred.train==labeled.data.lost.train$severe) # true rate
```

## Evaluation

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
options(scipen = 1, digits = 2)
require(dplyr)
require(ggplot2)
require(RcppRoll)
require(markovchain) # FIX: markovchain seems not compatible with R3.5.0
require(reshape)
require(survival)
require(caret)
require(e1071)
require(randomForest)
# knitr::opts_knit$set(root.dir = "~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/") # !! CHANGE PATH TO MATCH YOUR OWN
# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/prepare_violation.R")
# source("~/Documents/Projects/Github/CRIS/ViolationData/violation-data-analysis/Houston/src/roll_over.R")
#knitr::opts_knit$set(root.dir = "~/documents/projects/violation-data-analysis") # !! CHANGE PATH TO MATCH YOUR OWN

@@

num.days.lm <- lm(0*num.death+1*num.days.lost+1*num.days.restrict~last.quarter.lost+last.year.lost+last.three.years.lost
           +last.quarter.restrict+last.year.restrict+last.three.years.restrict
           +last.quarter.viol+last.year.viol+last.three.years.viol
           +last.quarter.death+last.year.death+last.three.years.death, 
           data=complete.active.quarters %>% filter(active))
summary(num.days.lm)
```
Linear Regression on num of deaths:

@@

#### K-means

@@

#### Spectral Clustering

Next, let's try out spectral clustering.  (TODO)

@@

# save the computationally costly svm model
#save(svm.model, file="./Houston/data/svm_radial.RData")
load("./Houston/output/svm_radial.RData")

tune_out <- tune.svm(severe~.,
                     data=labeled.data.lost.train,
                     gamma=10^(-3:3),
                     cost=c(0.01,0.1,1,10,100,1000) ) # e1071 tune functions have internal iteration limits, in this case, it is better to use another package

compareTable <- table (labeled.data.lost.train$severe, predict(svm.model))  # tabulate
mean(labeled.data.lost.train$severe != predict(svm)) # misclassification rate

pred.train <- predict(svm.model, labeled.data.lost.train)
mean(pred.train==labeled.data.lost.train$severe) # classification rate
```

It turns out SVM is not a good option for classifying highly skewed data. `tune.svm` does not work for this dataset due to its large size. Current SVM model with RBF kernel gives 100% misclassification rate.

### Random Forest

Random forest is self-bagging algorithm, which never overfits no matter what number of trees given. It is fast and performs well in most cases. We try to use random forests to classify the data.

```{r randomForest, include=FALSE}
# reformat labeled.data.lost.train$severe to factors so that `randomForest`
# can do a classification
labeled.data.lost.train$severe <- labeled.data.lost.train$severe %>% factor
labeled.data.lost.result$severe <- labeled.data.lost.result$severe %>% factor

# Note: nrow of `labled.data.lost.train` and `labeled.data.lost.result` are not equal
# so, we need to modify the longer one of them in order to run `predict` function.
# Let's do it on `labeled.data.lost.train`.
labeled.data.lost.train.trimmed <- labeled.data.lost.train[9:nrow(labeled.data.lost.train), ]

ptm <- proc.time() # start clock to time the process
rf.model <- randomForest(severe ~ last.quarter.lost+last.year.lost+last.three.years.lost
                 +last.quarter.restrict+last.year.restrict+last.three.years.restrict
                 +last.quarter.viol+last.year.viol+last.three.years.viol
                 +last.quarter.death+last.year.death+last.three.years.death,
                 data = labeled.data.lost.train.trimmed,
                 strata = mind_id,
                 nodesize = 240, # tuning parameter; does not affect much; may improve speed a little
                 ntree=200, # tuning parameter; no effect
                 mtry = 2 # tuning parameter; improves OOB error a little; see `rf.tune.mtry`
                 )
proc.time() - ptm # stop the clock

#save(rf.model, file="rf_model.RData") # base model, without any tuning
#save(rf.model, file="rf_model_tuned.RData") # tuned model

load("./Houston/output/rf_model.RData")
load("./Houston/output/rf_model_tuned.RData")

rf.tune.mtry <- tuneRF(labeled.data.lost.train[, c('last.quarter.lost','last.year.lost',
                                                'last.three.years.lost','last.quarter.restrict',
                                                'last.year.restrict', 'last.three.years.restrict',
                                                'last.quarter.viol','last.year.viol',
                                                'last.three.years.viol', 'last.quarter.death',
                                                'last.year.death', 'last.year.death',
                                                'last.three.years.death')],
                       labeled.data.lost.train[, 'severe']
                       )

#save(rf.model, file="rf_tune_mtry.RData") # mtry tuning results

pred <- predict(rf.model, data=labeled.data.lost.result)
freq <- table(pred, labeled.data.lost.result$severe)
sum(diag(freq))/sum(freq) # report accuracy

plot(rf.model) # plot the error rate vs number of trees

It turns out that random forest can classify the problem in a reasonable time, however, the false predication of true events is 0.88 after tuning. The OOB error is 1.1%. tuning does not improve the performance significantly.

Now, let's try an ensemble learning method, which is a combination of random forest and SVM.

### Ensemble Learning
